---
title: "Peer Assessment 1"
output: html_document
---
##Loading and preprocessing the data

```{r,message=FALSE}
library(dplyr)
library(ggplot2)
library(lubridate)
#setwd("c://temp/coursework-reproduceabledata/project1")
dat <- read.csv("activity.csv")
result <- dat %>% group_by(date) %>% summarize (total_steps = sum(steps,na.rm = TRUE ))
```

##What is mean total number of steps taken per day?

```{r}
g<- ggplot(result, aes(total_steps))
g+geom_histogram(binwidth = 3000)+ggtitle("Count of Total no of Steps Each Day")
mean_tot_steps <-mean(result$total_steps)
median_tot_steps <- median(result$total_steps)
```
mean total steps taken per day is `r  mean_tot_steps`

median total seps taken per day is `r median_tot_steps`

##What is the average daily activity pattern?

```{r}
result1 <- dat %>% group_by(interval) %>% summarize(avg_step = mean(steps, na.rm = TRUE))
plot(result1$interval, result1$avg_step, type="l", xlab = "interval", ylab = "avg_steps",main="Average daily activity pattern")
interval_of_max_avg_step <- result1[result1$avg_step == max(result1$avg_step),][1,1]

```
Interval with maximum no of avg steps is `r interval_of_max_avg_step`

##Imputing missing values 

Description of imputing strategy:


The strategy adopted for imputing missing values is to fill them with the corresponding interval average.
The missing values are separated out into a dataframe and then imputed with the interval averages using the 
interval averages calculated in the previous section. Then this data is combined with the rest of the data using rbind for calculating new totals and plotting.

```{r}
no_missing_values <- nrow(dat[is.na(dat$steps),])

```

no of missing values in the dataset is `r no_missing_values`

```{r}
ndata <- dat[is.na(dat$step),]
x <- vector()
for(n in c(1:2304)){x[n] <- result1[result1$interval== ndata[n,3],][1,2]}
xf <- data.frame(unlist(x))
result2 <- cbind(ndata,xf)
result2 <- mutate(result2, steps = unlist.x.)
result2 <- select(result2, steps:interval)
#append result2 to original dataframe sans the missing value rows 
new <- rbind(dat[!is.na(dat$steps),],result2)
#total number of steps taken each day
result3 <- new %>% group_by(date) %>% summarize (total_steps = sum(steps ))
#histogram for new dataframe with total steps per each day
g3<- ggplot(result3, aes(total_steps))
g3+geom_histogram(binwidth = 3000)+ggtitle("Count of Total no of Steps Each Day with NAs filled in")
mean_tot_steps_nafil <- mean(result3$total_steps)
median_tot_steps_nafil <- median(result3$total_steps)

```

mean total steps taken per day with NAs filled in  is `r sprintf("%.2f", mean_tot_steps_nafil)`

median total seps taken per day with NAs filled in  is `r sprintf("%.2f",median_tot_steps_nafil)`

###What is the impact of imputing missing data on the estimates of the total daily number of steps?

Mean and median values have increased after the NAs were filled in with interval averages. Also the median and mean values have become one and the same.

##Are there differences in activity patterns between weekdays and weekends?

```{r}
new1 <- mutate(new, day = ifelse(wday(new$date,label=TRUE) %in% c("Sat", "Sun"), "weekend", "weekday"))
result4 <- new1 %>% group_by(day,interval) %>% summarize(avg_steps = mean(steps))
library(lattice)
xyplot(avg_steps~interval|as.factor(day),data =result4,layout=c(1,2),type="l")

```

